### 2014-06-27 paper/section/conclusion.tex

\section{Conclusion and Future Work}

In this paper, we propose a I/O burst buffer architecture to burst I/O throughput, provide throughput-based, cost-based and queue write back model, and provide a simulation based on several benchmarks on TSUBAME supercomputer and AMAZON EC2 public cloud.
we use several nodes in each system as a I/O buffer nodes, and use data buffering to hide the low throughput between two systems connected by Internet.

From simulation result, we showed that our I/O burst buffer can increase I/O throughput when Internet throughput is far smaller than interconnection throughput as well as reducing the overall cost.

For future work, we plan to implement this I/O burst buffer architecture.