### 2014-07-01 paper/src/1_introduction.tex
0701

\section{Introduction}
\label{sec:introduction}
Supercomputers provide an increasing number of scientific applications with high computational power by a large number of processors, large bandwidth memory and interconnects.
%Data size need to be processed grows rapidly these days, which is known as Big Data, people perfer to use supercomputer to analyze Big Data for it supports high parallelism.
%Multi-user can use supercomputer at the same by using batch queue system which manage nodes and job submission.
%Although supercomputers can offer the high computational power, there are several situations where a supercomputer can not satisfy user's demands even there are still some idle nodes, consider there are 50 nodes available, a user submit a serial application use only one node for an hour, after that, a job using 50 nodes is submitted, and have to wait until previous job finished although there are 49 nodes remainning available.
%high parallel application runs faster on Supercomputer for fully usage of computing resource, but a serial application will occupy computing nodes for a long time causing other application must wait.
Although supercomputers can also offer high computational capacity, the computational resources are not unlimited. 
Supercomputers can not meet the demands of users when the demands exceed the limit. 
For example, on grand challenge projects, computational resources are exclusively used for the scientific discoveries~\cite{Scientific_Grand_Challenges_Crosscutting_Technologies_for_Computing_at_the_Exascale, Scientific_Grand_Challenges_in_National_Security_the_Role_of_Computing_at_the_Extreme_Scale}, and the system can not provide adequate resources for other jobs.
In addition, under power budget constrain~\cite{Exploring_Hardware_Overprovisioning_in_Power-constrained_High_Performance_Computing,Operation_of_TSUBAME_2_0_Green_Supercomputer_dealing_with_Power_Crisis}, a part of compute nodes are required to be shutdown to reduce its power consumption, which also leads inadequate of resources for all users.

One of solutions to provide adequate computational resources under the constrains, is federating the supercomputer with public clouds.
By moving a part of jobs to public clouds when there are not enough
compute nodes available, we can meet the needs for users' requests even under the constrains, which is known as \emph{cloud bursting}~\cite{Optimizing_Service_Level_Agreements_for_Autonomic_Cloud_Bursting_Schedulers, A_Framework_for_Data-Intensive_Computing_with_Cloud_Bursting}.
Although cloud bursting is technically feasible~\cite{Eucalyptus,stratos,Seagull}, 
and is employed by several private companies, there are several problems if we apply this technique to supercomputers.
One of the problems is a significant performance gap between supercomputers and public clouds especially in I/O performance
%there will still be several problems when we try to federate a supercomputer with a public cloud, 
%and there are several studies on cloud burst about cost\cite{Seagull}, execution time\cite{time_and_cost}, etc..
%Especially, low I/O performance between supercomputers and public clouds is critical.
%The biggest problem will be data transfer throughput between two environment, 
For example, if we migrate a part of jobs from a supercomputer to a public cloud or run jobs on a public cloud insted of a supercomputer, and the jobs need data located in a parallel file system (PFS), the data need to be transferred between the two system.
% supercomputers usual deal with Gigabyte or even Petabyte input and output of data, 
Because the two systems are usually geographically distributed, and network throughput between the systems is quite low,
the low I/O throughput suffers supercomputer users. Thus, improving I/O throughput between two systems, which are geographically distributed each other, is critical in federating supercomputers with public clouds.

In this paper, we propose I/O burst buffer architecture. The I/O burst buffers consists of I/O dedicated staging nodes, which cache hot files in the buffers, and enables asynchronous write back for improving both read and write operations to remote file systems.
We also model the I/O burst buffers to optimize  configurations of the I/O burst buffers according to dynamically changing environments
%\kento{Do you think this sentence is resonable to describe your model ?}.
According to our simulation, we can achieve a 4-20 times higher throughput (depends on data locality and other factors) by using 20 I/O buffer nodes on Amazon EC2 environment, as well as reduce 2-12 times of cost.
%\kento{please summirize the evaluation resutls here.}
%  architecture that uses multiple nodes in each system as an I/O nodes 
%to improve I/O throughput concurrent data transferring, and an I/O burst buffer model 
%that uses to switch between I/O burst buffer mode and direct connect mode.
%This I/O bursting buffer architecture can be mainly used two situation, first in data I/O, the other one will be when there are some nodes need to migrate from a cloud to another cloud with snapshot, snapshots can be transferred without be stored to shared storage.
%Since, I/O bursting buffer needs three times data transfer, sometimes direct transfer will achieve a higher throughput, I/O bursting buffer model using a evaluation throughput using I/O bursting buffer and without I/O bursting buffer to determine I/O bursting buffer mode or direct connection mode.
%Also, public cloud usually charges for nodes usage, I/O bursting buffers may reduce the computation time, 
%but I/O nodes will be charged for money, we also provide a cost-based model, to reduce the overall cost.

Our contributions can be summarized as following:
\begin{itemize}
	\item A I/O burst buffer architecture for increasing data transfer throughput between two systems;
	\item A throughput-based I/O burst buffer model, which estimate I/O throughput of systems with I/O burst buffers, and a cost-based I/O burst buffer model, which estimate overall cost give a system configuration;
        \item Evaluation of the I/O burst buffer architecture based the performance models by using real data obtained from several benchmarks from the TSUBAME supercomputer and Amazon EC2.
\end{itemize}
%2
The rest of this paper is organized as follows. 
In Section \ref{sec:motivation}, we clarify the motivation and the background.
We introduce an overview of the I/O bursting buffer architecture in Section \ref{sec:burst_buffer}, 
and show the performance models to evaluate the I/O bursting buffer architecture in Section \ref{sec:model}. 
In Section \ref{sec:evaluation}, we present our experimental results based of our performance models. %based on data obtained from several benchmark on TSUBAME V queue and Amazon EC2 
Finally, we detail related work in Section \ref{sec:related_work}, and conclusion in Section \ref{sec:conclusion}.