### 2014-07-01 paper/src/6_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper, we propose a I/O burst buffer architecture to burst I/O throughput, provide throughput-based, cost-based and queue write back model, and provide a simulation based on several benchmarks on TSUBAME supercomputer and AMAZON EC2 public cloud.
we use several nodes in each system as a I/O buffer nodes, and use data buffering to hide the low throughput between two systems connected by Internet.

From simulation result, we showed that our I/O burst buffer solution can burst I/O throughput about 4-20 times depends on cache hit rate as well as reduce overall cost about 2-12 times when Internet throughput is far smaller than interconnection throughput.

For future work, we plan to implement this I/O burst buffer architecture.